Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-09-29 forecast/paper/response-to-reviewers.tex

\begin{resp}
  We thank the editor and reviewers for their comments and the opportunity to make
  revisions. We have made a number of revisions, described in detail below.
  We believe these changes have strengthened the manuscript, and describe them in
  more detail below.
  We quote below the Reviewer comment and provide our response in this color.
\end{resp}

\section*{Reviewer 1}

  1. In their presentation of of evaluation metrics they do not discuss
  stability. For instance in [7] does the choice of the weights, using $\tau$ for
  $x\geq 0$ rather than some monotone function say some power $\tau^\alpha$ make a
  difference? Since as they indicate the measure is used for fitting as well as
  ascertainment, some stability investigation seems appropriate. I suspect here
  is no great effect but...

\begin{resp}
  The reviewer is correct: the choice of $\tau$ rather than a monotone function of
  $\tau$ will make a difference. But it seems that we should have been more clear
  in our description of WIS. The weights used here are those used to evaluate
  forecasts submitted to the COVID-19 Forecast Hub and the CDC. More broadly,
  this is a standard metric in the forecasting community (just as Mean Squared
  Forecast Error is standard), rather than an ``author decision''.  In principle,
  we would expect that different weights may alter the conclusions, though
  likely not substantively. Applying a monotone function to $\tau$ only effectively
  changes the quantile of interest, resulting in a mismatch between the coverage you
  tried to get and the coverage WIS is evaluating. Applying the function to $\tau$
  and $(1-\tau)$ results in an asymmetry, penalizing forecasts that miss on one
  side more than on the other.

@@

  3. This points to what is perhaps a more major issue. Can they identify in
  quantitative terms the consequences of the additional improvements in
  forecasting provided by these methods in terms of gains in resources available
  to deal with the epidemic? Unfortunately as they state, in the crucial periods
  of sharply rising cases, the additional benefit seems least.

\begin{resp}
  This is a good point, and we have made a few modifications described in more
  detail below (Reviewer 2, Point 3).
\end{resp}

  4. Despite the fact that a detailed description of terms in the text is
  impossible, it would I think be helpful to say if there is any significance to
  the names they associate with their first 2 and 4th indicators.

\begin{resp}
  It seems that the Reviewer is referring to CHNG-CLI, CHNG-COVID and DV-CLI. These are
  described in the Methods Section, Signals and Locations Subsection. To be more
  specific, Change 
  Healthcare is a large healthcare insurance claims processor. The difference
  between ``-CLI'' and ``-COVID'' is that the first measures insurance
  claims involving symptoms
  associated with COVID-like-illness  while the second requires a medical
  diagnosis (through 
  testing or presumed positive by the medical provider). The ``DV'' modifier
  corresponds to different claims processors (not Change Healthcare) who wish to
  remain anonymous.

@@

  We also have strong reasons for not using model-based approaches (one being the
  nonstationarity the Reviewer mentions above, but also the belief that these
  models are wrong). We felt that the related investigation in 2B was more
  pertinent.

  To respond more directly:

  $\bullet$ We have added a section to the Supplement (and point
  to it around 
  line 85 in the main text), that performs a Sign Test for differences in
  accuracy. P-values there tend to cluster near zero suggesting again that these
  improvements are statistically significant.

  $\bullet$ We also use a Diebold-Mariano test
  for comparing forecast accuracy. The results for this test are more mixed,
  likely due to nonstationarity the Reviewer mentions above (the indicators are
  more helpful in some periods than others, see also the upswing/downswing
  discussion in Section 2C). 
\end{resp}

@@

3. Are the results in this paper practically significant? I see no discussion of
this question, which seems unbelievable for me to be saying, given that the
authors refuse to consider statistical significance. And yet here I am.

\begin{resp}
  See also Reviewer 1 point 3. And the general comment below (Point 8).
  Practically, on average, if our goal is to predict 11 days ahead, when the AR
  model is at its best, CHNG-CLI
  doesn’t help (at least with these simple models). However, CHNG-COVID,
  DV-CLI, and Google-AA ``buy'' 4 extra days for the 
  same accuracy. CTIS-CLIIC ``buys'' about 5 extra days. What we mean here is
  that if we have a certain tolerance for inaccuracy, adding these signals increases the
  horizon at which we can predict within our tolerance. As for the
  practical significance of 4-5 extra days, this would of course depend on the use case.  However, cases grow
  exponentially, so making public policy changes 15 or 16 days ahead rather than
  11 can mean real gains. There also seems to be a fairly constant mapping
  between cases and future hospitalizations. Giving
  hospitals a few extra days to accumulate supplies and find personnel is
  meaningful. For predicting hotspots, we see similar results (gains of a few
  days) though with different indicators. 
\end{resp}

4. For the cumsum plots I mentioned above, and other related techniques, we can
plot all the model prediction error cumsum curves versus time and thereby
compare different models for size, nonstationarity and significance of
prediction improvement. The plots that have been presented in Figure 4 say do
none of those tasks.

\begin{resp}
  We added the cumulative-sum plot as described above. Additionally, the
  Reviewer's point aligns with much of  our motivation for Fig 4. If we were
  predicting a single location's 
  time series, then the cum-sum plot would provide all the information.
  However, we have a separate time series prediction problem for each HRR, and
  the upswings/downswings occur at different dates in different HRRs, so
  conditioning on forecast date for the comparison (as is done in a
  cumulative-sum plot) ends up adding together errors from different phases
  of the pandemic. In Figure 4, we instead condition on phase of the pandemic,
  which we think is more meaningful than calendar date. 
\end{resp}

@@

5. The idea of using Asof data is of course important but I have known about it
for twenty years and practiced it routinely; it has apparently been in routine
use in econometrics for 40 years.

\begin{resp}
  We appreciate the Reviewer's perspective on the ubiquity of Asof data in
  econometrics. But the epidemic modelling community is less aware. And it
  has a real and meaningful importance for model selection going forward. We
  hope that this paper will lead to greater appreciation of this issue in this
  community. 
\end{resp}

@@

\begin{resp}
  Fixed. Thank you for catching this.
\end{resp}

\subsection*{General Comments: }

8. In my understanding there are some standard models which have been relied
upon I guess millions of times during the pandemic. For example IHME's model.
The authors say a sentence or so about their choice of baseline, saying what
seems to me quite vague about the relation between their baseline and the
heavily used existing models. Wouldn't it be better to identify a specific
widely-used model explicitly as a baseline and then develop measures of
improvement over that specific widely used model? Also, might the authors
explain how the improvements observed by the authors stack up in terms of
concrete differences over the widely used models that people have been using.
Are there major misses that have taken place and could have been avoided?

\begin{resp}
  We have tried to focus mainly on simple time series models in this paper
  rather than comparing to other more elaborate models. A careful
  (out-of-sample) comparison of forecasting performance for state-of-the-art
  models can be found in reference 6 (Cramer et al.). That paper focuses on
  forecasting deaths rather than cases and at the state-level rather than
  disaggregated HRRs. However, their ranking (Figure 2 of Cramer et al.) puts the
  ``COVIDhub-Baseline'' about 10th out of 27 models. As we discuss in the
  manuscript, that baseline is  
  the
  same as the baseline we use for forecasting in this paper: it propagates today
  forward and resamples the residuals to create quantiles (we have expanded the
  description on page 6 to be more explicit). All our ``simplistic'' models beat
  this baseline (the AR(3) does as well). For comparison, IHME (along with 17 other
  ``state-of-the-art'' models) does not. We would argue that using any of the
  models in this paper would be preferable to relying on many of the
  ``widely-used'' models if the goal is forecast accuracy.

### 2021-09-29 forecast/paper/paper.tex
Pre-meeting modifications, responding to @jacobbien @huisaddison

While forecasting carries a rich literature that offers a wide range of
techniques, see e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to
very simple models, avoiding common enhancements such as order selection,
correction of outliers/anomalies in the data, and inclusion of regularization or
nonlinearities.
% Addition in response to Reviewer 1
For the same reason, we do not account for any environmental factors that may
well aid in forecasting---age structure, variants, holidays,
reporting patterns, vaccination, public health measures, etc.
% End addition
That said, analyses of forecasts submitted to the COVID-19
Forecast Hub \cite{ForecastHub} by a large community of modelers have shown that
simple, robust models have consistently been among the best-performing over the
pandemic, including time series models similar to those we
consider in what follows.
% Added in response to Reviewer 2
\cite{Cramer:2021} investigates the performance of submitted forecasts for
state-level deaths and shows that time series models perform among the top 5.
Our internal evaluations suggests that the same is true (at least, if not to a
greater extent) for case forecasting.

@@

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community who is sick, based on Delphi's COVID Trends
  and Impact Survey, in partnership with Facebook.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google search trends for anosmia and ageusia (Google-AA): A measure of  
  Google search volume for queries that relate to anosmia or ageusia (loss of
  smell or taste), based on Google's COVID-19 Search Trends data set.  
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

>>>>

While forecasting carries a rich literature that offers a wide range of
techniques, see e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to
very simple models, avoiding common enhancements such as order selection,
correction of outliers/anomalies in the data, and inclusion of regularization or
nonlinearities.
% Addition in response to Reviewer 1
For the same reason, we do not account for any environmental factors that may
well aid in forecasting---age structure, variants, holidays,
reporting patterns, vaccination, public health measures, etc.
% End addition
That said, analyses of forecasts submitted to the COVID-19
Forecast Hub \cite{ForecastHub} by a large community of modelers have shown that
simple, robust models have consistently been among the best-performing over the
pandemic, including time series models similar to those we
consider in what follows.
% Added in response to Reviewer 2
\cite{Cramer:2021} investigates the performance of submitted forecasts for
state-level deaths and shows that time series models perform among the top 5.
Our internal evaluations suggest that the same is true (at least, if not to a
greater extent) for case forecasting.

@@

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community who is sick, based on Delphi's COVID Trends
  and Impact Survey, in partnership with Facebook.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google search trends for anosmia and ageusia (Google-AA): A measure of  
  Google search volume for queries that relate to anosmia or ageusia (loss of
  smell or taste), based on Google's COVID-19 Search Trends data
  set~\cite{Wahltinez2020}.   
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

### 2021-09-29 forecast/paper/supplement.Rmd
Pre-meeting modifications, responding to @jacobbien @huisaddison

kableExtra::kable(
  bind_rows(dms) %>% 
    pivot_wider(names_from = forecaster, values_from = dm), 
  digits = 3, booktabs = TRUE, caption = "$P$-values for a Diebold-Mariano test for differences in forecast error. The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.")
```

@@

```{r cumulative-mean, fig.cap="This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows" }
cumulatives <- fcasts_honest %>%
  group_by(forecaster, forecast_date) %>%
  summarise(mw = Mean(wis),
            smw = Mean(strawman_wis),
            gmw = GeoMean(wis / strawman_wis)) %>%
  arrange(forecast_date) %>%
  mutate(`Cumulative Mean` = cummean(mw) / cummean(smw),
         `Cumulative Sum` = cumsum(mw) / cumsum(smw),
         `Cumulative Geo Mean` = exp(cummean(log(gmw))),
         `GM Regret` = cumsum(gmw) - cumsum(smw/smw),
         `AM Regret` = cumsum(mw) - cumsum(smw),
         `14 day trailing average` = RcppRoll::roll_meanl(mw / smw, n = 14L))

>>>>

kableExtra::kable(
  bind_rows(dms) %>% 
    pivot_wider(names_from = forecaster, values_from = dm), 
  digits = 3, booktabs = TRUE, caption = "$P$-values for a one-sided Diebold-Mariano test for improvement in forecast error. The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.")
```

@@

```{r cumulative-mean, fig.cap="This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows total U.S. cases as reported by JHU-CSSE." }
cumulatives <- fcasts_honest %>%
  group_by(forecaster, forecast_date) %>%
  summarise(mw = Mean(wis),
            smw = Mean(strawman_wis),
            gmw = GeoMean(wis / strawman_wis)) %>%
  arrange(forecast_date) %>%
  mutate(`Cumulative Mean` = cummean(mw) / cummean(smw),
         `Cumulative Sum` = cumsum(mw) / cumsum(smw),
         `Cumulative Geo Mean` = exp(cummean(log(gmw))),
         `GM Regret` = cumsum(gmw) - cumsum(smw/smw),
         `AM Regret` = cumsum(mw) - cumsum(smw),
         `14 day trailing average` = RcppRoll::roll_meanl(mw / smw, n = 14L))
